---
title: "STAT 340: Example Assignment Solution"
author: "Hintz, Erik (ehintz)"
output: pdf_document
---

\newcommand*{\T}{^{\top}}
\renewcommand*{\i}{\leftarrow}
\newcommand*{\isim}{\overset{\text{ind}}{\sim}}
\newcommand*{\deq}{\overset{\text{d}}{=}}
\newcommand*{\IN}{\mathbb{N}} 
\newcommand*{\IZ}{\mathbb{Z}}
\newcommand*{\IQ}{\mathbb{Q}}
\newcommand*{\IR}{\mathbb{R}}
\newcommand*{\IC}{\mathbb{C}}
\newcommand*{\Geo}{\operatorname{Geo}}
\newcommand*{\Exp}{\operatorname{Exp}}
\newcommand*{\Poi}{\operatorname{Poi}}
\newcommand*{\NVM}{\operatorname{NVM}}
\newcommand*{\Par}{\operatorname{Par}}
\newcommand*{\IG}{\operatorname{IG}}
\newcommand*{\LN}{\operatorname{LN}}
\newcommand*{\Cauchy}{\operatorname{Cauchy}}
\newcommand*{\Log}{\operatorname{Log}}
\renewcommand*{\U}{\operatorname{U}}
\newcommand*{\B}{\operatorname{B}}
\newcommand*{\Bin}{\operatorname{Bin}}
\newcommand*{\Bern}{\operatorname{Bern}}
\newcommand*{\Beta}{\operatorname{Beta}}
\newcommand*{\NB}{\operatorname{NB}}
\newcommand*{\N}{\operatorname{N}}
\newcommand*{\I}{\mathbbm{1}}
\newcommand*{\rd}{\mathrm{d}} 
\renewcommand*{\mod}{\operatorname{mod}}
\newcommand*{\arginf}{\operatorname*{arginf}}
\newcommand*{\argsup}{\operatorname*{argsup}}
\newcommand*{\ran}{\operatorname{ran}}
\newcommand*{\rank}{\operatorname{rank}}
\newcommand*{\sign}{\operatorname{sign}}
\newcommand*{\round}{\operatorname{round}}
\renewcommand*{\L}{\mathcal{L}}
\renewcommand*{\Re}{\operatorname{Re}}
\renewcommand*{\Im}{\operatorname{Im}}
\newcommand*{\Li}{\operatorname*{Li}}
\renewcommand*{\P}{\mathbb{P}} 
\newcommand*{\E}{\mathbb{E}} 
\newcommand*{\med}{\operatorname{med}}
\newcommand*{\Var}{\operatorname{Var}}
\newcommand*{\cov}{\operatorname{cov}}
\newcommand*{\cor}{\operatorname{Cor}}
\newcommand*{\R}{\textsf{R}}
\newcommand*{\eps}{\varepsilon}
\renewcommand*{\th}{\mathbf{\theta}}
\newcommand*{\ba}{\mathbf{a}}
\newcommand*{\bnu}{\mathbf{\nu}}
\newcommand*{\bb}{\mathbf{b}}
\newcommand*{\be}{\mathbf{e}}
\newcommand*{\blambda}{\mathbf{\lambda}}
\newcommand*{\bzero}{\mathbf{0}}
\newcommand*{\bone}{\mathbf{1}}
\newcommand*{\bt}{\mathbf{t}}
\newcommand*{\bx}{\mathbf{x}}
\newcommand*{\by}{\mathbf{y}}
\newcommand*{\bz}{\mathbf{z}}
\newcommand*{\bu}{\mathbf{u}}
\newcommand*{\bU}{\mathbf{U}}
\newcommand*{\bv}{\mathbf{v}}
\newcommand*{\bw}{\mathbf{w}}
\newcommand*{\btheta}{\mathbf{\theta}}
\newcommand*{\bS}{\mathbf{S}}
\newcommand*{\bT}{\mathbf{T}}
\newcommand*{\bZ}{\mathbf{Z}}
\newcommand*{\bY}{\mathbf{Y}}
\newcommand*{\bX}{\mathbf{X}}
\newcommand*{\bmu}{\mathbf{\mu}}
\newcommand*{\hmu}{\hat{\mu}}
\newcommand*{\hsigma}{\hat{\sigma}}
\newcommand*{\hmuMC}{\hat{\mu}^\text{MC}}


```{r setup, include=FALSE}
## Put commands you do not want to be printed here 
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
doPDF <- FALSE # logical if pdfs of plots shall be produced 
```

## Question 1

Since $X\sim\N(0,1)$, we know that the density of $X$ is given by

$$ \phi(x) = \frac{1}{\sqrt{2\pi}} \exp\left(-\frac{x^2}{2}\right),\quad x\in\IR.$$

We can use the \R\ command `rnorm()` to sample $n=2000$ realizations
$X_1,\dots,X_n \isim \N(0,1)$. 

```{r}
set.seed(2) # for reproducibility
n <- 2000 # sample size
x <- rnorm(n) # n-vector containing iid realizations of N(0,1)
est <- c(mean(x), var(x))
names(est) <- c("Estimated mean", "Estimated sd")
print(est)
```

Thus, we find that $\hat{\mu}$ = `r round(est[1], 2)` and $\hsigma$ = `r round(est[2], 2)`.

\newpage

## Question 2 

Next, we plot a histogram of the sample and overlay the true standard-normal
density. Aside: if `doPDF = TRUE`, compiling this document produces `.pdf` 
files of the plots (which can then easily be included in a Latex document,
for instance). 

```{r, fig.align = "center", fig.width = 7, fig.height = 5, fig.show = "hold"}
if(doPDF) pdf(file = (file <- "fig_histogram_density.pdf"),
              width = 7, height = 5)
hist(x, freq = FALSE, main = "Histogram of 10,000 iid N(0,1) realizations")
curve(dnorm(x), col = 2, lty = 2, lwd = 2, add = TRUE)
if(doPDF) dev.off() # don't forget to turn off the 'pdf device' 
```

