# ActSc 632 Assignment 2 Solutions
##### Spring 2023
##### Department of Statistics and Actuarial Science, University of Waterloo

## Question 1
Determine the number of predictors in this data set, and whether they are quantitative of qualita-
tive. How many observations are there? How many observations are classiﬁed as a “bad credit”?
“good credit”?

```{r}
library(MASS)
library(CASdatasets)

# Load the data
data(credit)

# Summary of the data
summary(credit)
print(c(sum(credit$class == 1),
        sum(credit$class == 0)))

```
There are 20 predictors. Duration, credit amount, age are quantitative, and the other
are qualitative. Depending on the encoding used for the data set (which diﬀers from source
to source) the predictors installment rate, residence since, existing credits and num dependents
may be encoded as integers and thus be a priori be considered quantitative. But when looking
at the meaning of the diﬀerent values these predictors can take, it seems best to treat them as
categorical. It won’t matter in our analysis though, as these 4 predictors do not end up being
considered in our models. and the others are categorical. There are 300 observations out of 1000
classiﬁed as bad credit.


```{r}

#box plots
attach(credit)
# pdf("box1.pdf")
par(mfrow=c(1,2))
boxplot(age[class==0],age[class==1],names=c("Good","Bad"),ylab = "age",col=c("blue","orange"))
boxplot(duration[class==0],duration[class==1],names=c("Good","Bad"),ylab = "duration",col=c("blue","orange"))
dev.off()
detach(credit)

#q2 -- plots of two predictors asked
attach(credit)
par(mfrow=c(1,1))
csG<-savings[class==0]
csB<-savings[class==1]
dG<-duration[ class==0]
dB<-duration[class==1]
# pdf("dur-sav.pdf")
plot(dG,csG,col=c("green"),pch=0,axes = F)
par(new=T)
plot(dB,csB,col=c("red"),pch=1,xlab="duration",ylab="savings",axes=T)
dev.off()
detach(credit)

attach(credit)
# pdf("dur-credh.pdf")
par(mfrow=c(1,1))
chG<-credit_history[class==0]
chB<-credit_history[class==1]

plot(dG,chG,col=c("green"),pch=0,axes = F)
par(new=T)
plot(dB,chB,col=c("red"),pch=1,xlab="duration",ylab="credit_history",axes=T)
dev.off()
detach(credit)
```