# ActSc 632 Assignment 1 Solutions
##### Spring 2023
##### Department of Statistics and Actuarial Science, University of Waterloo

```{r}
# Install the packages if necessary
# install.packages("insuranceData")
# install.packages("doParallel")
# install.packages("dplyr")
# install.packages("tibble")

library(insuranceData)
library(foreach)
library(dplyr)
library(tibble)

options(tibble.print_max = Inf)
```

## Question 1
Download the data set dataOhlsson and briefly discuss the data found in there (e.g., how many
rating factors, what are the levels for each, how is the exposure determined, etc.). If there are
any problems with the data, explain how you dealt with them.

#### Solution:

As stated in the R documentation for this dataset:

> The data for this case study comes from the former Swedish insurance company Wasa,
and concerns partial casco insurance, for motorcycles this time. It contains aggregated
data on all insurance policies and claims during 1994-1998; the reason for using this
rather old data set is confidentiality; more recent data for ongoing business can not be
disclosed.

```{r}
# Load the data
data(dataOhlsson)

# change categorical columns into factors
dataOhlsson <- within(dataOhlsson, {
    zon <- factor(zon)
    mcklass <- factor(mcklass)
    bonuskl <- factor(bonuskl)
    kon <- factor(kon)})

# regroup the column bonuskl into 3 levels
levels(dataOhlsson$bonuskl) <- list("1" = c("1","2"),
                                    "2" = c("3","4"),
                                    "3" = c("5","6","7"))  

# regroup the column mcklass into 3 levels
# For details, visit: https://stackoverflow.com/questions/13559076/group-numeric-values-by-the-intervals
dataOhlsson$motorage_gr <- cut(dataOhlsson$fordald, 
                               breaks = c(0, 2, 5, 100), 
                               labels = c("1", "2", "3"), 
                               right = FALSE)

summary(dataOhlsson)
```

In this data set we have 6 rating factors given by agarald (age of driver),which goes from 0 to
99, kon (sex of driver), either M (male) or K (female), zon (geographical zone, going from 1 to 7,
generally going from more to less urban), mcklass (class of the vehicle, with 7 possible classes, as
determined by the EV ratio of engine power to vehicle weight), fordald (age of vehicle, a numerical
value between 0 and 99), bonuskl (bonus class based on experience, going from 1 to 7: a new
driver starts with bonus class 1; for each claim-free year the bonus class is increased by 1; after
the first claim the bonus is decreased by 2; the driver can not return to class 7 with less than 6
consecutive claim free years). There are 64548 observations in this data set, with one for each
driver observed. We see that for some observations the duration is 0. The exposure (or duration)
is determined using policy-years, i.e., how long the contract was in effect for each driver. For
each observation we also have the number of claims (antskad) and the claim cost (skadkost).

## Question 2
Use only the rating factors contained in the current tariff. Compute the exposure (in policy
years), the claim frequency and the average claim severity for each rating factor (i.e., for each
possible value of each rating factor).

#### Solution:
```{r}
motorcycle <- data.frame(rating.factor = c(rep("Zone", nlevels(dataOhlsson$zon)),
                                           rep("Vehicle class", nlevels(dataOhlsson$mcklass)),
                                           rep("Vehicle age", nlevels(dataOhlsson$motorage_gr)),
                                           rep("Bonus class", nlevels(dataOhlsson$bonuskl))),
                         class = c(levels(dataOhlsson$zon),
                                   levels(dataOhlsson$mcklass),
                                   levels(dataOhlsson$motorage_gr),
                                   levels(dataOhlsson$bonuskl)),
                         stringsAsFactors = FALSE)

new.cols <- foreach (rating.factor = c("zon","mcklass","motorage_gr","bonuskl"), .combine = rbind) %do% {
                nclaims <- tapply(dataOhlsson$antskad, dataOhlsson[[rating.factor]], sum)
                sums <- tapply(dataOhlsson$duration, dataOhlsson[[rating.factor]],sum)
                severity <- tapply(dataOhlsson$skadkost,dataOhlsson[[rating.factor]],sum)
                n.levels <- nlevels(dataOhlsson[[rating.factor]])
                contrasts(dataOhlsson[[rating.factor]]) <-
                    contr.treatment(n.levels)[rank(-sums, ties.method = "first"), ]
                data.frame(duration = sums, n.claims = nclaims, totcostperm = severity/1000)}

motorcycle <- cbind(motorcycle, new.cols)
rm(new.cols)

```
After grouping the classes for the vehicle age into 3 groups (0-2, 3-5, 6 and up) and
the bonus class into three groups (1-2,3-4,5-7) we get
```{r}
motorcycle
```

## Question 3
Use a relative Poisson glm to determine relativities for the claim frequency, using the current
rating factors. Provide a 95% confidence interval for each relativity. Comment on the overall fit
of this model to the data.

#### Solution:
```{r}
# Group the data by rating factors
grDataOhlsson <- dataOhlsson %>% group_by(bonuskl, zon, mcklass, motorage_gr) %>% summarize(duration = sum(duration), 
                                                                                            antskad=sum(antskad),
                                                                                            skadkost=sum(skadkost))

# Fit the Poisson GLM
model.freq <- glm(antskad ~ zon + mcklass + motorage_gr + bonuskl + offset(log(duration)), 
                  family = poisson, 
                  data = grDataOhlsson[grDataOhlsson$duration>0,])

# Examine model fit
with(model.freq, cbind(res.deviance = deviance, 
                       df = df.residual, 
                       p = pchisq(deviance, df.residual, lower.tail = FALSE)))
```

With a deviance of $360.2168$ on $389$ degrees of freedom, the model seems a reasonable fit (p-value of $0.8495$).

```{r}
# Generate the table for relative frequency
P.zon = contrasts(dataOhlsson$zon)
P.mcklass = contrasts(dataOhlsson$mcklass)
P.motorage_gr = contrasts(dataOhlsson$motorage_gr)
P.bonuskl = contrasts(dataOhlsson$bonuskl)

print(P.zon)
print(P.mcklass)
print(P.motorage_gr)
print(P.bonuskl)

mat.freq = summary(model.freq)$coefficients[, 1:2]

table.q3 = matrix(0, nrow = dim(motorcycle)[1] + 1, ncol = 5)
colnames(table.q3) = c("Rating factor", "Class", "Multiplier", "Lower bound", "Upper bound")


# Intercept
estimate = mat.freq[1, 1]
std.error = mat.freq[1, 2]
table.q3[1, ] = c("Intercept", "0", exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Zone
estimate = P.zon %*% mat.freq[2:7, 1]
std.error = P.zon %*% mat.freq[2:7, 2]
table.q3[2:8, ] = cbind(rep("Zone", 7), seq(7), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Vehicle class
estimate = P.mcklass %*% mat.freq[8:13, 1]
std.error = P.mcklass %*% mat.freq[8:13, 2]
table.q3[9:15, ] = cbind(rep("Vehicle class", 7), seq(7), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Vehicle age
estimate = P.motorage_gr %*% mat.freq[14:15, 1]
std.error = P.motorage_gr %*% mat.freq[14:15, 2]
table.q3[16:18, ] = cbind(rep("Vehicle age", 3), seq(3), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Bonus class
estimate = P.motorage_gr %*% mat.freq[16:17, 1]
std.error = P.motorage_gr %*% mat.freq[16:17, 2]
table.q3[19:21, ] = cbind(rep("Bonus class", 3), seq(3), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

as_tibble(table.q3)
```

Note that I also accepted answers based on the quasi-Poisson family where a dispersion parameter is estimated, causing the CIs to be different from the above.

## Question 4
Use a Gamma glm (with log link function) to determine relativities for the severity, still using
the current rating factors. Provide a 95% confidence interval for each relativity. Comment on
the overall fit of this model to the data.

#### Solution:
```{r}  

model.sev <- glm(skadkost/antskad ~ zon + mcklass + motorage_gr + bonuskl, 
                 family=Gamma("log"), 
                 data=grDataOhlsson[grDataOhlsson$skadkost>0,],
                 weights=antskad)

with(model.sev, cbind(res.deviance = deviance, 
                      df = df.residual, 
                      p = pchisq(deviance, df.residual, lower.tail = FALSE)))
```

The severity model doesn’t seem to be a very good fit. We reject the hypothesis that the data comes from this model based on the residual deviance, given by $351$ on $164$ degrees of freedom, as its corresponding p-value is $1.13842 \times 10^{-15}$.

```{r}
# Generate the table for relative severity
mat.sev = summary(model.sev)$coefficients[, 1:2]

table.q4 = matrix(0, nrow = dim(motorcycle)[1] + 1, ncol = 5)
colnames(table.q4) = c("Rating factor", "Class", "Multiplier", "Lower bound", "Upper bound")

# Intercept
estimate = mat.sev[1, 1]
std.error = mat.sev[1, 2]
table.q4[1, ] = c("Intercept", "0", exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Zone
estimate = P.zon %*% mat.sev[2:7, 1]
std.error = P.zon %*% mat.sev[2:7, 2]
table.q4[2:8, ] = cbind(rep("Zone", 7), seq(7), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Vehicle class
estimate = P.mcklass %*% mat.sev[8:13, 1]
std.error = P.mcklass %*% mat.sev[8:13, 2]
table.q4[9:15, ] = cbind(rep("Vehicle class", 7), seq(7), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Vehicle age
estimate = P.motorage_gr %*% mat.sev[14:15, 1]
std.error = P.motorage_gr %*% mat.sev[14:15, 2]
table.q4[16:18, ] = cbind(rep("Vehicle age", 3), seq(3), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

# Bonus class
estimate = P.motorage_gr %*% mat.sev[16:17, 1]
std.error = P.motorage_gr %*% mat.sev[16:17, 2]
table.q4[19:21, ] = cbind(rep("Bonus class", 3), seq(3), exp(estimate), exp(estimate - 1.96 * std.error), exp(estimate + 1.96 * std.error))

as_tibble(table.q4)
```


## Question 5
Assess whether rating factors for the policyholder’s age and sex would have a significant impact.
Include an interaction term. You should group the age factor into intervals of 0-30 and 30-100. 

#### Solution:
```{r}
dataOhlsson$driverage_gr <- cut(dataOhlsson$agarald, 
                                breaks = c(0, 30,100), 
                                labels = c("1", "2"), 
                                right = FALSE)

motorcycle.new <-
    data.frame(rating.factor = c(rep("Bonus class", nlevels(dataOhlsson$bonuskl)),
                                 rep("Vehicle class", nlevels(dataOhlsson$mcklass)),
                                 rep("Vehicle age", nlevels(dataOhlsson$motorage_gr)),
                                 rep("Zone", nlevels(dataOhlsson$zon)),
                                 rep("Driver age", nlevels(dataOhlsson$driverage_gr)),
                                 rep("Sex", nlevels(dataOhlsson$kon))),
               class = c(levels(dataOhlsson$bonuskl),
                         levels(dataOhlsson$mcklass),
                         levels(dataOhlsson$motorage_gr),
                         levels(dataOhlsson$zon),
                         levels(dataOhlsson$driverage_gr),
                         levels(dataOhlsson$kon)),
               stringsAsFactors = FALSE)

new.cols <- foreach (rating.factor = c("zon", "mcklass", "motorage_gr", "bonuskl", "driverage_gr", "kon"), .combine = rbind) %do% {
    nclaims <- tapply(dataOhlsson$antskad,dataOhlsson[[rating.factor]], sum)
    sums <- tapply(dataOhlsson$duration, dataOhlsson[[rating.factor]],sum)
    severity <- tapply(dataOhlsson$skadkost,dataOhlsson[[rating.factor]],sum)
    n.levels <- nlevels(dataOhlsson[[rating.factor]])
    contrasts(dataOhlsson[[rating.factor]]) <-
        contr.treatment(n.levels)[rank(-sums, ties.method = "first"), ]
    data.frame(duration = sums, n.claims = nclaims, sev = severity/nclaims)}

motorcycle.new <- cbind(motorcycle.new, new.cols)
rm(new.cols)

grDataOhlsson.new <- dataOhlsson %>% group_by(bonuskl, zon, mcklass, motorage_gr, kon, driverage_gr) %>% summarize(duration = sum(duration),
                                                                                                                   antskad=sum(antskad),
                                                                                                                   skadkost=sum(skadkost))

# Fit the Poisson GLM for frequency
model.freq.new <- glm(antskad ~ zon + mcklass + motorage_gr + bonuskl + kon * driverage_gr +offset(log(duration)),
                     family = poisson, 
                     data = grDataOhlsson.new[grDataOhlsson.new$duration>0,])

with(model.freq.new, 
     cbind(res.deviance = deviance, df = df.residual, 
     p = pchisq(deviance, df.residual, lower.tail = FALSE)))


pchisq(model.freq.new$deviance - model.freq$deviance,
       model.freq.new$df.residual - model.freq$df.residual, 
       lower.tail=FALSE)
```
We have combined the ages into 2 groups given by the breaks 0, 30, 100. When including age and sex with an interaction term, for the frequency the deviance goes to $742.76$ on $1277$ degrees of freedom. The LRT statistic we get $742.76 - 360.22 =382.55$ on $1277 - 389=888$ degrees of freedom. The corresponding p-value is $1$, suggesting that the simpler model is a better fit. 

```{r}
# Fit the Gamma GLM for severity
model.sev.new <- glm(skadkost/antskad ~ zon + mcklass + motorage_gr + bonuskl + kon * driverage_gr, 
                     family = Gamma("log"), 
                     data = grDataOhlsson.new[grDataOhlsson.new$antskad>0,],
                     weights = antskad)

with(model.sev.new, 
     cbind(res.deviance = deviance, df = df.residual, 
     p = pchisq(deviance, df.residual, lower.tail = FALSE)))

pchisq(model.sev.new$deviance - model.sev$deviance * summary(model.sev)$dispersion / summary(model.sev.new)$dispersion, 
       model.sev.new$df.residual - model.sev$df.residual,
       lower.tail=FALSE)
```

For the severity however, the LRT statistic is given by $253.12$ on $124$ degrees of freedom, with corresponding p-value given by $7.09 \times 10^{−11}$. Hence in this case, the age and sex do appear to provide a model that has a better fit. We also see that for both the frequency and the severity models, while sex is not a significant factor, the interaction term between age and sex and the coefficient for the age group 0-30 are both significant.

## Question 6

Now combine your multiplier estimates for the frequency and severity data to get multipliers (and associated 95% confidence intervals) for the premium overall and then propose a new tariff based on your analysis. Compare the results to the old tariff.

#### Solution:
```{r}


# Generate the table for new tariff
table.q6 = matrix(0, nrow = dim(motorcycle)[1] + 1, ncol = 5)
colnames(table.q6) = c("Rating factor", "Class", "Multiplier", "Lower bound", "Upper bound")

# Intercept
estimate.freq = mat.freq[1, 1]
std.error.freq = mat.freq[1, 2]
estimate.sev = mat.sev[1, 1]
std.error.sev = mat.sev[1, 2]
table.q6[1, ] = c("Intercept", "0", exp(estimate.freq + estimate.sev), 
                  exp(estimate.freq + estimate.sev - 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)), 
                  exp(estimate.freq + estimate.sev + 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)))

# Zone
estimate.freq = P.zon %*% mat.freq[2:7, 1]
std.error.freq = P.zon %*% mat.freq[2:7, 2]
estimate.sev = P.zon %*% mat.sev[2:7, 1]
std.error.sev = P.zon %*% mat.sev[2:7, 2]
table.q6[2:8, ] = cbind(rep("Zone", 7), seq(7), exp(estimate.freq + estimate.sev), 
                  exp(estimate.freq + estimate.sev - 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)), 
                  exp(estimate.freq + estimate.sev + 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)))

# Vehicle class
estimate.freq = P.mcklass %*% mat.freq[8:13, 1]
std.error.freq = P.mcklass %*% mat.freq[8:13, 2]
estimate.sev = P.mcklass %*% mat.sev[8:13, 1]
std.error.sev = P.mcklass %*% mat.sev[8:13, 2]
table.q6[9:15, ] = cbind(rep("Vehicle class", 7), seq(7), exp(estimate.freq + estimate.sev), 
                   exp(estimate.freq + estimate.sev - 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)), 
                   exp(estimate.freq + estimate.sev + 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)))
# Vehicle age
estimate.freq = P.motorage_gr %*% mat.freq[14:15, 1]
std.error.freq = P.motorage_gr %*% mat.freq[14:15, 2]
estimate.sev = P.motorage_gr %*% mat.sev[14:15, 1]
std.error.sev = P.motorage_gr %*% mat.sev[14:15, 2]
table.q6[16:18, ] = cbind(rep("Vehicle age", 3), seq(3), exp(estimate.freq + estimate.sev), 
                    exp(estimate.freq + estimate.sev - 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)), 
                    exp(estimate.freq + estimate.sev + 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)))
# Bonus class
estimate.freq = P.motorage_gr %*% mat.freq[16:17, 1]
std.error.freq = P.motorage_gr %*% mat.freq[16:17, 2]
estimate.sev = P.motorage_gr %*% mat.sev[16:17, 1]
std.error.sev = P.motorage_gr %*% mat.sev[16:17, 2]
table.q6[19:21, ] = cbind(rep("Bonus class", 3), seq(3), exp(estimate.freq + estimate.sev), 
                    exp(estimate.freq + estimate.sev - 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)), 
                    exp(estimate.freq + estimate.sev + 1.96 * sqrt(std.error.freq^2 + std.error.sev^2)))

# Append the old tariff
tariff.old <- c(0, 
                7.768, 4.227, 1.336, 1.000, 1.734, 1.402, 1.402,
                0.625, 0.769, 1.000, 1.406, 1.875, 4.062, 6.873,
                2.000, 1.200, 1.000,
                1.250, 1.125, 1.000)

table.q6 <- cbind(table.q6, tariff.old)
as_tibble(table.q6)
```
We see that the new tariff is much higher than the old one for Vehicle Age 1 and 2. We also
see that for Zone 5 to 7, the new tariff suggests to lower the premium compared to the baseline
of Zone 4, while for the old tariff they were all higher than the baseline. The number of claims
for these 3 zones is very small though, so it seems like it might be best to not make such an
important change based on such a small sample.

## Question 7
Comment on any further analysis that should be considered before deciding on a final tariff.

#### Solution:
we saw that, especially for the severity, the age and sex seem to be significant factors
that should be considered. On the other hand, looking at the results for coefficients in each of the
frequency and severity models (reproduced below), we see that Zones 5,6,7 (note that although
the results below are based on the ordering of levels by decreasing order of duration, and therefore
do not necessarily correspond to the original numbering, for these 3 particular classes are actually
the same as in the original ordering) do not seem to be significant, and should therefore probably
be combined with the baseline of Zone 4. Finally, we should explore the use of models other than
the gamma distribution for the severity, to see if a better fit can be obtained.
